\vspace{-0ex}Bench: Evaluating Arbitrary Workloads

### 2018-02-24 paper/submission.tex
Added citation for Distributed systems.
Cleaned up spelling mistakes in paper..

\begin{abstract}
Software applications use medium like HDD (Hard Disk Drives) and SSD (Solid State Drives) to store information and data. The storage workloads for these applications vary, ranging from highly sequntial workloads to ones which utilize the drives in a random order. Choosing the correct medium of storage is essential to delivering optimal performance for these applications. Hence, it is crucial to benchmark these workloads to identify the most suitable storage medium. Furthermore, it is important in general to recognize which workloads suit a specific storage medium the most. The work presented in this paper introduces an application "bench" which provides a way to benchmark these varying types of workloads using a consumer/producer architecture. Using \textit{bench}, we evaluate and present the result of some common industrial workloads and their performance metrics on a variety of storage medium.
\end{abstract}

@@

To be able to reduce or even elminate this bottleneck, it is critical to realize which class of storage media technology performs most efficiently for a given workload. Using this information, one can make better decisions with regards to system design and provide a more definitive performance analysis for the overall system. However, all algorithms produce different workloads and in most cases, a software application itself employs many different algorithms. Hence, it makes more sense to refer to a workload as part of an application rather than that of an algorithm. This abstraction is also extendable to systems which employ different sets of applications to provide some service, an example being an operating system such as Unix. In such a case, workload may refer to the overall workload of the operating system rather than the seperate workloads of the applications within it.

Keeping this in mind, it is important to understand that different storage media performs differently for particular types of workloads. Conventional rotating media like hard disk drives perform optimally in a sequential workload, whereas flash based storage such as used in solid state drives can perform admirably even in a random-heavy workload. However, flash based storage costs more than rotating media storage per giga byte. This means that an engineer designing a system needs to make careful decisions on the inclusion of hard drives, solid state drives and even a combination of both. Driving these decisions is usually a benchmark, highlighting the performance and cost characteristics of the different types of storage media. However, performing such an exhaustive benchmark is a non-trivial task. Different classes of applications produce completely different workloads. For example, a web page may produce small-random oriented workloads whereas a paging algorithm for an operating system might produce a long-sequential one.

@@

All the platters in a disk drive are organised as a set of concentric circles, each of which is known as a \textbf{track}. These tracks are then further divided into smaller regions called \textbf{sectors}, and the sector size denotes the granularity of I/O for any given disk drive. The tracks on a platter are packed together incredibly tightly, with hundreds of tracks fitting within the width of a human hair.

@@

Flash memory groups a set of these cells into \textbf{pages}. A single page is the granularity for reading and writing to flash memory, and in general, page size is set to 4 KB. One problem with flash memory is that a page cannot be written to unless it has been previously erased. Unfortunately, flash memory performs the erase operation on a \textbf{block} granularity. A block is a collection of pages and usually block size is much greater than page size. The reason that this is the case is because the erase operation consumes a lot of voltage, which in turn consumes more power and at the same time damages the structure of the transistor in a cell. Hence, an erase operation batches a number of pages together and erases them all at the same time. While this improves the endurance and performance of flash memory, the assymetric granularity of writes and erases create an interesting problem: updates to flash memory cannot be done in-place.

@@

\section{Design}
We have designed \textit{bench} with keeping this background in mind. In this section, we explain the design choices made for \textit{bench}. We argue that these design choices allow \textit{bench} to benchmark most if not all types of workloads. By using a probability distribution for tasks and an exponential distribution for arrival time, along with a simulated distributed environment, \textit{bench} is theoretically capable of modelling most industrial workloads.

@@

An I/O task is essentially a tuple which describes an I/O operation of a certain size at a specific offset in the drive. Depending upon the relation of this I/O task to previous I/O tasks, it may be classified as either random or sequential. Random I/O is usally smaller in size and in most cases bears no correspondance to previously issued I/O tasks in the system. Sequential I/O on the other hand is usually larger and has a direct correspondance to previously issued tasks. For example, consider a workload which reads a drive at 1 MB per read request. If the first read for the workload takes place at offset 0, then the next sequential read will take place at offset 1 MB. On the other hand, if the workload processed reads randomly, then such a correspondance between the two read requests wouldn't likely exist. While this assumption is not always true, it is a fair one to make.

Based on this primitive of an I/O task, we can break down any given workload into a list of I/O tasks it performs. Using this list, we can build a distribution which describes the probability of each type of task occuring. This distribution is simple to build as there are only four types of I/O tasks: random read, random write, sequential read and sequential write. If an application was to process this list, then it would execute the same workload as the original application, albeit without any application context. While not having context generally makes no difference to most workloads, it makes accurately benchmarking certain workloads difficult.

Consider a workload which processes I/O tasks with a very strict order. A random read in this workload is followed by a sequential write, which is then followed by a random read and so on. While it is possible to benchmark such a workload through a probability distribution, a benchmark based solely on task distribution will be inaccurate as it will not be able to guarentee the order of the tasks. Future versions of \textit{bench} aim to find a workaround.

@@

A \textbf{Poisson point process} is a process in which events occur continously and independently at a constant average rate \cite{kingman1993poisson}. This means that as time $t \Rightarrow \infty$, the average of the intervals between events in a point process converge. Hence, for a system with a known average, \textit{such as a workload}, the poisson point process can model the interval times between a pair of events. An \textbf{exponential distribution} is the probability distribution which describes an interval between a pair of events in a poisson point process \cite{buzen1973computational}. Hence, using an exponential distribution, we can accurately model the arrival of work for any given workload.

However, this approach is not perfect. Suppose a workload experiences different arrival times under different curcumstances, such as time of day. A workload which experiences a different arrival time of clients between day and night cannot be modelled correctly with an average. The average shadows the different stress levels which the workload produces and coalesces them into a single, more normalized arrival time. This can lead to inaccurate benchmarks. We aim to solve this issue in future revisions of \textit{bench}.

\subsection{Distributed Architecture}
Most industrial applications are built as distributed systems which divide work amongst different entities in order to better process requests from clients. In a generic distributed application, a frontend system collects a request from a user and forwards it to a backend system. This backend system processes this request and ships the information back to the frontend or to the client.

@@

As the network is not a bottleneck anymore, we decided that we can combine the frontend and the backend into a single application and simulate them using threads instead of building two separate applications, one for the frontend and one for the backend, and then connect them together using a network simulator. We claim that having two separate applications because of something which as is does not limit performance provides no benefit to having an application which simulates entities using threads. Infact, we argue that bulding separate applications leads to more bug prone software and will overall provide deprecating results.

@@

Work profiles store the probability distribution of the tasks using a cumulative distribution sum and intervals within this sum represent the task. For exmaple, consider a workload with the following distribution: random read (50\%), random write (10\%), sequential read (15\%) and sequntial write (25\%). The work profile will store these as: random read ([1, 50]), random write ([51, 60]), sequential read ([61, 75]) and sequntial write ([76, 100]). Now, to generate a task with respect to the distribution, all we need to do is generate a uniformly distributed random number $x$ in [1, 100]. Depending on the interval bracket $x$ falls in, we choose a size and an offset for the I/O task.

@@

The current implementation of \textit{bench} uses \texttt{rand()} from the standard library to generate random numbers. We are aware of the fact that \texttt{rand()} has shown to \textbf{NOT} be uniformly distributed. For future versions, we are looking into using the linux system call \texttt{getrandom()} as our entropy source.

@@

\begin{enumerate}
    \item Acquire an exponential variate of the arrival averge in microseconds. Let this be $n$.
    \item Call \texttt{usleep} with $n$ as the parameter. 
    \item Upon waking up, generate an I/O task and append it to the circular queue shared with the consumer. 
    \item Repeat. 
\end{enumerate}

The consumer on the other hand does a lot more. Its workflow is divided into separate phases which are represented using an \texttt{enum}. They are: \texttt{initialize}, \texttt{process} and \texttt{output}. The initialize phase is for setting up allocations and opening the required file descriptor. The process phase is a loop which processes I/O tasks acquired from the circular queue. Finally, the output phase is where the consumer dumps all statistically relevant information for making the performance graphs.

@@

Consider a situation where the circular queue is empty and the consumer goes to sleep, hoping the producer will wake it up. If the timer thread cancels the prodcuer before it can wake the consumer up, then the consumer sleeps indefinitely and the only way to come back to a workable state is to manually kill the program. To avoid such a situation, the following ordering policy takes place:

@@

We also evaluated standard benchmarks to corrobate the performance characteristics of chosen media. Hence, we chose a sequential based workload for the HDD and sequential + random read based workload for the SSD.

@@

\section{Conclusion}
It is possible to build a benchmarking application capable of benchmarking most if not all workloads. The corroboration of the result with the expectation means that \textit{bench} is a viable option for performing becnhmarks on workloads.

>>>>

\begin{abstract}
Software applications use medium like HDD (Hard Disk Drives) and SSD (Solid State Drives) to store information and data. The storage workloads for these applications vary, ranging from highly sequential workloads to ones which utilize the drives in a random order. Choosing the correct medium of storage is essential to delivering optimal performance for these applications. Hence, it is crucial to benchmark these workloads to identify the most suitable storage medium. Furthermore, it is important in general to recognize which workloads suit a specific storage medium the most. The work presented in this paper introduces an application "bench" which provides a way to benchmark these varying types of workloads using a consumer/producer architecture. Using \textit{bench}, we evaluate and present the result of some common industrial workloads and their performance metrics on a variety of storage medium.
\end{abstract}

@@

To be able to reduce or even eliminate this bottleneck, it is critical to realize which class of storage media technology performs most efficiently for a given workload. Using this information, one can make better decisions with regards to system design and provide a more definitive performance analysis for the overall system. However, all algorithms produce different workloads and in most cases, a software application itself employs many different algorithms. Hence, it makes more sense to refer to a workload as part of an application rather than that of an algorithm. This abstraction is also extendable to systems which employ different sets of applications to provide some service, an example being an operating system such as Unix. In such a case, workload may refer to the overall workload of the operating system rather than the separate workloads of the applications within it.

Keeping this in mind, it is important to understand that different storage media performs differently for particular types of workloads. Conventional rotating media like hard disk drives perform optimally in a sequential workload, whereas flash based storage such as used in solid state drives can perform admirably even in a random-heavy workload. However, flash based storage costs more than rotating media storage per gigabyte. This means that an engineer designing a system needs to make careful decisions on the inclusion of hard drives, solid state drives and even a combination of both. Driving these decisions is usually a benchmark, highlighting the performance and cost characteristics of the different types of storage media. However, performing such an exhaustive benchmark is a non-trivial task. Different classes of applications produce completely different workloads. For example, a web page may produce small-random oriented workloads whereas a paging algorithm for an operating system might produce a long-sequential one.

@@

All the platters in a disk drive are organized as a set of concentric circles, each of which is known as a \textbf{track}. These tracks are then further divided into smaller regions called \textbf{sectors}, and the sector size denotes the granularity of I/O for any given disk drive. The tracks on a platter are packed together incredibly tightly, with hundreds of tracks fitting within the width of a human hair.

@@

Flash memory groups a set of these cells into \textbf{pages}. A single page is the granularity for reading and writing to flash memory, and in general, page size is set to 4 KB. One problem with flash memory is that a page cannot be written to unless it has been previously erased. Unfortunately, flash memory performs the erase operation on a \textbf{block} granularity. A block is a collection of pages and usually block size is much greater than page size. The reason that this is the case is because the erase operation consumes a lot of voltage, which in turn consumes more power and at the same time damages the structure of the transistor in a cell. Hence, an erase operation batches a number of pages together and erases them all at the same time. While this improves the endurance and performance of flash memory, the asymmetric granularity of writes and erases create an interesting problem: updates to flash memory cannot be done in-place.

@@

\section{Design}
We have designed \textit{bench} with keeping this background in mind. In this section, we explain the design choices made for \textit{bench}. We argue that these design choices allow \textit{bench} to benchmark most if not all types of workloads. By using a probability distribution for tasks and an exponential distribution for arrival time, along with a simulated distributed environment, \textit{bench} is theoretically capable of modeling most industrial workloads.

@@

An I/O task is essentially a tuple which describes an I/O operation of a certain size at a specific offset in the drive. Depending upon the relation of this I/O task to previous I/O tasks, it may be classified as either random or sequential. Random I/O is usually smaller in size and in most cases bears no correspondence to previously issued I/O tasks in the system. Sequential I/O on the other hand is usually larger and has a direct correspondence to previously issued tasks. For example, consider a workload which reads a drive at 1 MB per read request. If the first read for the workload takes place at offset 0, then the next sequential read will take place at offset 1 MB. On the other hand, if the workload processed reads randomly, then such a correspondence between the two read requests wouldn't likely exist. While this assumption is not always true, it is a fair one to make.

Based on this primitive of an I/O task, we can break down any given workload into a list of I/O tasks it performs. Using this list, we can build a distribution which describes the probability of each type of task occurring. This distribution is simple to build as there are only four types of I/O tasks: random read, random write, sequential read and sequential write. If an application was to process this list, then it would execute the same workload as the original application, albeit without any application context. While not having context generally makes no difference to most workloads, it makes accurately benchmarking certain workloads difficult.

Consider a workload which processes I/O tasks with a very strict order. A random read in this workload is followed by a sequential write, which is then followed by a random read and so on. While it is possible to benchmark such a workload through a probability distribution, a benchmark based solely on task distribution will be inaccurate as it will not be able to guarantee the order of the tasks. Future versions of \textit{bench} aim to find a workaround.

@@

A \textbf{Poisson point process} is a process in which events occur continuously and independently at a constant average rate \cite{kingman1993poisson}. This means that as time $t \Rightarrow \infty$, the average of the intervals between events in a point process converge. Hence, for a system with a known average, \textit{such as a workload}, the Poisson Point Process can model the interval times between a pair of events. An \textbf{exponential distribution} is the probability distribution which describes an interval between a pair of events in a Poisson Point Process \cite{buzen1973computational}. Hence, using an exponential distribution, we can accurately model the arrival of work for any given workload.

However, this approach is not perfect. Suppose a workload experiences different arrival times under different circumstances, such as time of day. A workload which experiences a different arrival time of clients between day and night cannot be modeled correctly with an average. The average shadows the different stress levels which the workload produces and coalesces them into a single, more normalized arrival time. This can lead to inaccurate benchmarks. We aim to solve this issue in future revisions of \textit{bench}.

\subsection{Distributed Architecture}
Most industrial applications are built as distributed systems which divide work among different entities in order to better process requests from clients. In a generic distributed application, a frontend system collects a request from a user and forwards it to a backend system. This backend system processes this request and ships the information back to the frontend or to the client \cite{tanenbaum2007distributed}.

@@

As the network is not a bottleneck anymore, we decided that we can combine the frontend and the backend into a single application and simulate them using threads instead of building two separate applications, one for the frontend and one for the backend, and then connect them together using a network simulator. We claim that having two separate applications because of something which as is does not limit performance provides no benefit to having an application which simulates entities using threads. In fact, we argue that building separate applications leads to more bug prone software and will overall provide deprecating results.

@@

Work profiles store the probability distribution of the tasks using a cumulative distribution sum and intervals within this sum represent the task. For example, consider a workload with the following distribution: random read (50\%), random write (10\%), sequential read (15\%) and sequential write (25\%). The work profile will store these as: random read ([1, 50]), random write ([51, 60]), sequential read ([61, 75]) and sequential write ([76, 100]). Now, to generate a task with respect to the distribution, all we need to do is generate a uniformly distributed random number $x$ in [1, 100]. Depending on the interval bracket $x$ falls in, we choose a size and an offset for the I/O task.

@@

The current implementation of \textit{bench} uses \texttt{rand()} from the standard library to generate random numbers. We are aware of the fact that \texttt{rand()} has shown to \textbf{NOT} be uniformly distributed. For future versions, we are looking into using the Linux system call \texttt{getrandom()} as our entropy source.

@@

\begin{enumerate}
    \item Acquire an exponential variate of the arrival average in microseconds. Let this be $n$.
    \item Call \texttt{usleep} with $n$ as the parameter. 
    \item Upon waking up, generate an I/O task and append it to the circular queue shared with the consumer. 
    \item Repeat. 
\end{enumerate}

The consumer on the other hand does a lot more. Its work flow is divided into separate phases which are represented using an \texttt{enum}. They are: \texttt{initialize}, \texttt{process} and \texttt{output}. The initialize phase is for setting up allocations and opening the required file descriptor. The process phase is a loop which processes I/O tasks acquired from the circular queue. Finally, the output phase is where the consumer dumps all statistically relevant information for making the performance graphs.

@@

Consider a situation where the circular queue is empty and the consumer goes to sleep, hoping the producer will wake it up. If the timer thread cancels the producer before it can wake the consumer up, then the consumer sleeps indefinitely and the only way to come back to a workable state is to manually kill the program. To avoid such a situation, the following ordering policy takes place:

@@

We also evaluated standard benchmarks to corroborate the performance characteristics of chosen media. Hence, we chose a sequential based workload for the HDD and sequential + random read based workload for the SSD.

@@

\section{Conclusion}
It is possible to build a benchmarking application capable of benchmarking most if not all workloads. The corroboration of the result with the expectation means that \textit{bench} is a viable option for performing benchmarks on workloads.
